\textls[-0]{\thetitle}

### 2023-03-18 metatheory-condensed.tex
W-types

>>>>

For each pair of types identifying a constructor, the first type may be thought of
as the non-inductive part of the constructor, and the second type as the arity
of the inductive part.
For example, the signature~$\wsig_{\texttt{tm}}$ of the W-type modeling
the inductive type~\lsti{tm} (\cref{fig:stlc-mechanized})
is constructed as follows,
where $\bbzero$ is the bottom type~$\bot$, $\bbone$ the unit type~$\top$, $\bbtwo$ the boolean type,
and $T_{\texttt{id}}$ a type encoding~\lsti{id}:
%
\begin{align*}
\small
\begin{tabular}{@{}llll@{}}
\lsti!tm_unit : tm!
&
\lsti!tm_var : id -> tm!
&
\lsti!tm_abs : id -> tm -> tm!
&
\lsti!tm_app : tm -> tm -> tm!
\smallskip\\
$\wsig_{\texttt{tm}}^0 \coloneq \WSigAdd{\WSigEmp}{\bbone}{\lambda \_ . \bbzero}$ 
& $\wsig_{\texttt{tm}}^1 \coloneq \WSigAdd{\wsig_{\texttt{tm}}^0}{T_{\texttt{id}}}{\lambda \_ . \bbzero}$
& $\wsig_{\texttt{tm}}^2 \coloneq \WSigAdd{\wsig_{\texttt{tm}}^1}{T_{\texttt{id}}}{\lambda \_ .\bbone} $ 
& $\wsig_{\texttt{tm}} \coloneq \WSigAdd{\wsig_{\texttt{tm}}^2}{\bbone}{\lambda \_ .\bbtwo}$
\end{tabular}
\end{align*}

\noindent
Thus, the encoding of \lsti{tm_abs} has type
$T_{\texttt{id}} \to (\bbone \to \El{\wcode{\wsig_{\texttt{tm}}}}) \to \El{\wcode{\wsig_{\texttt{tm}}}}$,
and that of \lsti{tm_app} has type
$\bbone \to (\bbtwo \to \El{\wcode{\wsig_{\texttt{tm}}}}) \to \El{\wcode{\wsig_{\texttt{tm}}}}$.
These types are strictly positive by construction.

%We illustrate the structure of W-type signature (as a list) here.
%$\wsig_{\texttt{tm}}^0, \wsig_{\texttt{tm}}^1$ are (the subparts of) the
%signatures of constructors \{"tm\_unit"\} and \{"tm\_unit","tm\_app"\}
%respectively. Each item in this W-type signature, having two types, corresponds
%to one constructor, where the first type stores the non-recursive part of the
%constructor and the second type stores the ``arity'' of the  recursive part. For
%example, the type of $"app" : ("tm" \to "tm" \to "tm")$ $\cong (\bbone \times
%(\bbtwo \to "tm") \to "tm")$, corresponds to the premise of rule
%\ruleref{tm/wsup} (where $\wsigproj{0}{1}{\wsig_{\texttt{tm}}^1} = \bbone,
%\wsigproj{0}{2}{\wsig_{\texttt{tm}}^1} = \lambda \_ . \bbtwo$). This type means
%there is no information on non-recursive part (thus use unit type $\bbone$) and
%the ``arity'' of the recursive part is $\bbtwo$ and thus $\wsig_{\texttt{tm}}^1
%= \WSigAdd{\wsig_{\texttt{tm}}^0}{\bbone}{\lambda \_ .\bbtwo}$.  The story is
%similar for $\wsig_{\texttt{tm}}^0$, where there is no recursive part at all and
%thus the ``arity'' is zero.\YZ{Feels like it's more straightforward to just list
%the encodings of the four constructors of tm in a table, assuming an encoding of
%the identifier type. And this paragraph should be moved to the earlier paragraph
%on W-types. I'll take a stab at this.}
%
%Generally speaking, the expressiveness of W-types comes from the ``arity'', as
%it encodes the case when recursive structure stores as the output of a function,
%mirroring the strict positivity of Inductive type.
